---
title: "Assignment 3"
author: "Uliana Plotnikova"
date: "September 14, 2019"
output: html_document
---


```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"

```


```{r}
library(stringr)
```

### Excersize 3 
### Copy the introductory example. The vector name stores the extracted names.

```{r}
name<-unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name

```
#### 3.1. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name

```{r}
FirstName<-str_extract(unlist(str_extract(name,"[:alpha:]{2,} |, [:print:]{2,}")),"[A-Z](.+?)[a-z]+")
FirstName

```

```{r}
LastName<-str_extract(unlist(str_extract(name,"[A-z] [:alpha:]{2,}[a-z]$|[:print:]{2,},")),"[A-z][a-z]+|[A-z][a-z]")
LastName
```


```{r}
Full_Name<- str_c(FirstName,LastName , sep= " "  )
data.frame (Full_Name)
```



#### 3.2. Construct a logical vector indicating wether a character has a title (i.e, Rev. and Dr.)


```{r}
title<-unlist(str_match_all(name, "Dr.|Rev."))
title
```

#### 3.3  Construct a logical vector indicating whether a character has a second name

```{r}
sec_name<-unlist(str_match_all(name," [A-Z]\\."))
sec_name
```



### Excersize 4.
#### Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.
 
```{r}
 ## a.[0-9] equivalent to [:digit:] or \d: digits, 0 1 2 3 4 5 6 7 8 9.The string "[0-9]+\\$" means that there are one or more numbers followed by $ sign. 
string1="[0-9]+\\$"
example1 = c("123456$", "123$")
str_detect(example1, string1)
```


```{r} 
 ## b.  \b is useful for word edge and can help to be more specific in the location of matches.
##The string below helps to find the word that has any letters a to z and no more than 4 letters
string2="\\b[a-z]{1,4}\\b"
example2=c("gym", "bug", "ball")
str_detect(example2, string2)
```


```{r}
 ## c. Finds pattern that ends with .txt
string3=".*?\\.txt$"
example3=c("book.txt", "bear56.txt")
str_detect(example3, string3)
```


```{r}
 ##  d.Finds pattern of numbers in the format 00/00/0000 
string4="\\d{2}/\\d{2}/\\d{4}"
example4=c("08/26/2019", "09/03/2019 School time")
str_detect(example4, string4)
```


```{r}
 ## e. Finds tags in an HTML documents
string5="<(.+?)>.+?</\\1"
example5=c('Weather<tag>is<tag>beautiful', '<tag>I love school</tag>')
str_detect(example4, string4)
```




### Exersize 5
#### Rewrite the expression [0-9]+\\$ in a way that all elements are altered but the expression performs the same task

```{r}
case1="[0-9]+\\$"
example=c("7$","123$")
str_detect(example,case1)
case2="\\d+[$]"
example=c("7$","123$")
str_detect(example,case2)
```


### Excersize 9

```{r}
code <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
```



```{r}
unlist(str_extract_all(code, "[:alpha:]"))
```

```{r}
code1 <- unlist(str_extract_all(code, "[[:upper:].!]"))
code1
```


```{r}
code2=paste(code1, collapse="")
Secret_message <- c( str_replace_all(code2, "[\\.]",  "  "))
Secret_message

```

