---
title: "Assignment 7"
author: "Uliana Plotnikova"
date: "October 13, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Library preparation

```{r}
library(XML)
library(RJSONIO)
library(DT)

```


Books files were manually created and saved in different formats on local device and also can be found on: 

1.https://raw.githubusercontent.com/uplotnik/Data607/master/Books.html

2.https://raw.githubusercontent.com/uplotnik/Data607/master/Books.xml

3.https://raw.githubusercontent.com/uplotnik/Data607/master/Books.json

# HTML file

```{r}
html_parsed<-htmlParse(file = "file:///C:/Users/a/Desktop/607/Books.html")

html_table<-readHTMLTable(html_parsed, stringsAsFactors = FALSE)
html_table<-html_table[[1]]


datatable(html_table)

```

```{r}
str(html_table)
```

# XML file

```{r}
xml_books <- "file:///C:/Users/a/Desktop/607/Books.xml"
book = xmlParse(xml_books)
xml_table <- xmlToDataFrame(book)
datatable(xml_table)
```

```{r}
str(xml_table)
```

# JSON file

```{r}
json_books<-fromJSON(content = "file:///C:/Users/a/Desktop/607/Books.JSON")

#create data frame.
json_table<-do.call("rbind", lapply(json_books[[1]], data.frame, stringsAsFactors = FALSE))

datatable(json_table)

```
```{r}
str(json_table)
```

# Conclusion

All three files saved in different formats look identical when we view data frames using library(DT).
However, variables in HTML and JSON files are saved as Characters (except "Pages" variable which is a Number vector in JSON), while all variables in XML file are Factors.





