---
title: "Data 607 Project 2b"
author: "Uliana Plotnikova"
date: "October 06, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    
---


# Extract and prepare data

```{r warning=FALSE}
library ("dplyr")
library ("tidyr")
library(data.table)
library(ggplot2)
library(kableExtra)

```



```{r}
url<-"https://raw.githubusercontent.com/uplotnik/Data607/master/Airline-Safety.csv"
a<- read.csv(url)
kable(head(a, 10))
```
###Rename columns
```{r}
##Rename columns for easy manipulation 
dataset <- a %>%
  rename("Incidents. 1985-1999"= incidents_85_99,
         "Fatal accidents. 1985-1999" = fatal_accidents_85_99,
         "Fatalities. 1985-1999"= fatalities_85_99,
         "Incidents. 2000-2014" = incidents_00_14,
         "Fatal accidents. 2000-2014" = fatal_accidents_00_14,
        "Fatalities. 2000-2014" = fatalities_00_14)
head(dataset,5)
```
###Gather columns

```{r}
data2 <- dataset %>%
  gather(Episode, Quantity, 'Incidents. 1985-1999' :'Fatalities. 2000-2014') %>%
  arrange(airline)
 head(data2,5) 
```

### Separate column


```{r}
data3<-data2 %>% 
  separate(Episode, into = c("Episode", "YEAR"), sep = "\\. " )
head(data3,5)
```

### Sum Episodes
```{r}
tx<-aggregate(data3$Quantity, by=list(Episode=data3$Episode, Year=data3$YEAR), FUN=sum) %>%
  group_by(Episode,Year) %>% 
  summarise(count=(x)) %>% 
  mutate(perc=count/sum(count))
 ##rename column
tx<- dplyr::rename(tx, 'Total'='count', 'Percent'='perc')
head (tx,5 )
```




### Total Episodes by Years

```{r}
p<-ggplot(tx, aes(x=Episode, y=Total, group=Year)) +
  geom_line(aes(color=Year))+
  geom_point(aes(color=Year))
p
```


```{r}
total<-aggregate(data3$Quantity, by=list(Years=data3$YEAR), FUN=sum) %>%
  mutate(pct_change = ((x/lag(x) - 1) * 100))

 ##rename column
total<- dplyr::rename(total, 'Total'='x', 'Percentage change'='pct_change')
total
```


The total quantity of episodes dropped on almost 51% after 1999.



```{r}

    ggplot(tx, aes(x = Episode, y = Percent, fill = Year, label = scales::percent(Percent))) + 
    geom_col(position = 'dodge') + 
    geom_text(position = position_dodge(width = .9),    
              vjust = -0.5,    
              size = 3) + 
    scale_y_continuous(labels = scales::percent)+  theme(axis.text=element_text(angle=90))+  ggtitle("Total Episodes by Years")



```



# Analyse flight episodes 1985-1999
```{r}
newdata <- data3 %>% 
  filter(YEAR=="1985-1999") 
  head(newdata,5)
```

### The safest airlines in 1985-1999	

```{r}
group_by(newdata, airline)  %>% filter(Quantity == min(Quantity)) %>% arrange(Quantity)  %>% select (airline,Episode,Quantity)

```


```{r}
newdata %>%
  group_by(Episode, YEAR) %>%
  summarise(
    mean = mean(Quantity), max=max(Quantity))%>%
  arrange(Episode)

```



### Total episodes by airlines

```{r}
total<-aggregate(newdata$Quantity, by=list(Carrier=newdata$airline, Seats=newdata$avail_seat_km_per_week), FUN=sum)
 ##rename column
total<- dplyr::rename(total, 'Total'='x')%>%
 arrange(desc(Total))
head (total,15 )
```






### Visualization
```{r}
g2<-ggplot(data=total, aes(x=Carrier, y=Total, group=10)) +
  geom_line(arrow = arrow())+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs( x="Carrier", y="Total")+ggtitle("Total Episodes 1985-1999")
g2
```


```{r}

ggplot(newdata, aes(airline, Quantity, color = Episode)) +
  geom_point()+  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

# Analyse fliht episodes 2000-2014


```{r}
newdata1 <- data3 %>% 
  filter(YEAR=="2000-2014") 
  head(newdata1,5)
```



```{r}
newdata1 %>%
  group_by(Episode, YEAR) %>%
  summarise(
    mean = mean(Quantity), max=max(Quantity))%>%
  arrange(Episode)

```

### The safest airlines 2000-2014
```{r}
group_by(newdata1, airline)  %>% filter(Quantity == min(Quantity)) %>% arrange(airline)  %>% select (airline,Episode,Quantity)

```



### Total episodes by airlines

```{r}
total1<-aggregate(newdata1$Quantity, by=list(Carrier=newdata$airline, Seats=newdata$avail_seat_km_per_week), FUN=sum)
 ##rename column
total1<- dplyr::rename(total1, 'Total'='x')%>%
  arrange(desc(Total))
head (total1,15 )
```

### Visualization


```{r}
g2<-ggplot(data=total1, aes(x=Carrier, y=Total, group=10)) +
  geom_line(arrow = arrow())+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs( x="Carrier", y="Total")+ggtitle("Total Episodes 2000-2014")
g2

```
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+


```{r}

ggplot(newdata1, aes(airline, Quantity, color = Episode)) +
  geom_point()+  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```







